---
title: "Miami Dade Analysis"
author: "Madison Volpe"
date: "1/18/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(stringi)
setwd("/Users/madisonvolpe/Desktop/Miami_Analysis")
books <- read.csv("JailBookins_1.19.csv")
```

# Pick Appropriate Charges

```{r}
# 322340010A - DWLS (No Prior Felony)
# 32234002B -  DWLS/2ND OFFENSE
# 32234002C - DWLS/3RD & SUBS OFFN
# 32420101  - DWLS/FIN RESP/REFUSE
# 322340010B2 - DWLS/FIN SUSP/SUB V (subsequent violation)
# 322340010B1 - DWLS/FINANCIAL SUSP
# 32234005    - DWLS/HABITUAL
# 32234002A   - DWLS/KNOWINGLY
```

```{r}
books$Charge1 <- as.character(books$Charge1) 
books$Charge2 <- as.character(books$Charge2) 
books$Charge3 <- as.character(books$Charge3)

charges <- data.frame(charges = sort(books$Charge1)) 
charges2 <- data.frame(charges = sort(books$Charge2)) 
charges3 <- data.frame(charges = sort(books$Charge3))

charges <- rbind(charges,charges2,charges3)
charges <- charges[!duplicated(charges$charges),]

rel_charges <- c("DWLS-NO PRIOR FELONY", "DWLS/2ND OFFENSE", "DWLS/3RD & SUBS OFFN", "DWLS/FIN RESP/REFUSE",
                 "DWLS/FIN SUSP/SUB V", "DWLS/FINANCIAL SUSP", "DWLS/HABITUAL", "DWLS/KNOWINGLY")

rm(charges, charges2, charges3)
```

# Filter Appropriate Charges 

```{r}
books_fil <- books %>% filter(Charge1 %in% rel_charges|
                                 Charge2 %in% rel_charges|
                                  Charge3 %in% rel_charges)
```

# Naive Analysis

```{r}
#Average Length of Stay (ALOS) Over the Last 180 Days: 33.20. Likewise, the MDCR reports that most of their prison population is concentrated in LOS of less than or equal to one month. 

# The FTP Ratio is computed based on data from the Florida DHMSV sent by Linda Foster. According to those number for calendar year 2917 (CY2017) -- there were 309484, total sanctions in Miami-Dade of these 16,574 were suspended due to "Fail to Pay Ct Financial Obligation" (that is about 5%). I pplied this 5% ratio to the total number of DWLS suspensions for each CY year. Therefore, for example (4050) * .05 is 203. The following calculations are range estimates for these people. If we assume defendants stayed in prison 1 day, then multiply 230 * 203 to get a total one.day.cost of 46,690 for the year. If they stayed two days we get a total two.day.cost of 93,380 (2*230*203) for the year. Finally, the average.stay represents multiplying the ALOS of stay in Miami-Dade Corrections and Rehabilitation, which is less than or equal to 30 days (230*30*203)-- thereby leading to a cost  of 1,400,700. 

books_fil$BookDate <- as.character(books_fil$BookDate)
books_fil$BookDate <- lubridate::mdy(books_fil$BookDate)
books_fil$year <- year(books_fil$BookDate)

books_fil %>%
  group_by(year) %>%
  summarise(n=n(),FTP.Ratio = ceiling((n*.05)), one.day.cost = 230*FTP.Ratio,
  two.day.cost = (2*230)*FTP.Ratio, average.stay= (30*230)*FTP.Ratio)
  
```

# Examine Charge Codes 

```{r}
ChargeCodes <-books_fil$ChargeCode1[match(rel_charges, books_fil$Charge1)] 
ChargeCodes<- unique(as.character(ChargeCodes))
```

# Creation of Charge Categories

```{r}
# Charge Code 1 
Charge1 <- data.frame(code = books_fil$ChargeCode1, charge = books_fil$Charge1)
Charge1$codes <- str_sub(Charge1$code, start = 1, end = 3)

# Charge Code 2 
Charge2 <- data.frame(code = books_fil$ChargeCode2, charge = books_fil$Charge2)
Charge2$codes <- str_sub(Charge2$code, start = 1, end = 3)

# Charge Code 3 
Charge3 <- data.frame(code = books_fil$ChargeCode3, charge = books_fil$Charge3)
Charge3$codes <- str_sub(Charge3$code, start = 1, end =3)

Charge1 <-Charge1[!duplicated(Charge1),]
Charge2 <- Charge2[!duplicated(Charge2),]
Charge3 <- Charge3[!duplicated(Charge3),]

Charges <- rbind(Charge1, Charge2, Charge3)
Charges <- Charges[!duplicated(Charges),]

rm(Charge1, Charge2, Charge3)

Charges <- Charges %>%
            arrange(codes)

names(Charges)[1] <- "Statute"
Charges$Statute_orginal <- Charges$Statute
Charges$Statute <- gsub("00$", "", Charges$Statute)
Charges$Statute <- gsub("0$", "", Charges$Statute)
Charges$Statute <- gsub("^0", "", Charges$Statute)
```

```{r}
# read in scraped offenses
official_statues <-read.csv("MiamiOffenses.csv")  
official_statues <- official_statues[-1]

# get rid of punctuation 
official_statues$Statute<-gsub("[().-]","",official_statues$Statute)
```

```{r}
# join charges from bookings data with official statues
Charges_Joined<- left_join(Charges, official_statues)
```

```{r}
#More cleaning by Charge Code 

Charges_Joined <- Charges_Joined %>%
  mutate(CodeLength = nchar(Statute)) %>%
  mutate(Zero_Consecutive = str_count(Statute, "000")) %>%
  mutate(Zero_Consecutive_end = str_count(Statute, "0{2}(?=\\w{1,4}$)")) %>%
  mutate(ChargeType = str_extract(Statute,"C$|D$|F$|I$|J$|L$|M$|N$|T$|X$|Z$|A$|G$|B$|H$"))
  


Charges_Joined <- Charges_Joined %>%
  mutate(Statute = ifelse(Zero_Consecutive == 1 & is.na(Desc), gsub("(?<=0{1})0{2}", "", Statute, perl = T),
                          Statute))%>%
  mutate(Statute = ifelse(Zero_Consecutive_end == 2 & is.na(Desc), gsub("0{2}(?=\\w{1}$)","",Statute, perl = T),
                          Statute)) %>%
  mutate(Statute = ifelse(Zero_Consecutive_end == 1 & codes != 320 & is.na(Desc),
                          gsub("0{2}(?=\\w{1,4}$)","",Statute, perl = T),
                          Statute)) %>%
  mutate(Statute = ifelse(Zero_Consecutive == 0 & Zero_Consecutive_end == 0 & is.na(Desc) & !codes %in%
                            c(316,320,322,784,790,796,806,810,812,893), gsub("0", "", Statute), Statute)) %>%
  mutate(Statute = ifelse(codes == 316 & is.na(Desc) & grepl("0{1}(?=\\w{1,4}$)", Statute, perl = T),
                          gsub("0{1}(?=\\w{1,4}$)", "", Statute, perl = T),
                          Statute))%>%
  mutate(Statute = ifelse(codes == 322 & is.na(Desc) & grepl("0{1}(?=\\w+\\D{1}$)", Statute, perl = T),
                        gsub("0{1}(?=\\w+\\D{1}$)", "", Statute, perl = T),
                        Statute)) %>%
  mutate(Statute = ifelse(is.na(Desc) & is.na(ChargeType) & codes == 893,
                         str_remove(Statute,"C$|D$|F$|I$|J$|L$|M$|N$|T$|X$|Z$|A$|G$|B$|H$"),
                          Statute)) %>%
  select(Statute, charge, codes, Statute_orginal)
```

```{r}
Charges_Joined <- left_join(Charges_Joined, official_statues)
Charges_Joined <- Charges_Joined[-c(1:131),]
sum(is.na(Charges_Joined$Desc)) 
check <- Charges_Joined[is.na(Charges_Joined$Desc),]
```

```{r}
Charges_Joined <- Charges_Joined %>%
  mutate(Statute = ifelse(codes == 812 & is.na(Desc),gsub("0{1}(?=\\w{1,3}$)", "", Statute, perl = T),
                          Statute))%>%
  select(Statute, charge, codes, Statute_orginal)
```

```{r}
Charges_Joined <- left_join(Charges_Joined, official_statues)
sum(is.na(Charges_Joined$Desc)) #86 not bad
check <- Charges_Joined[is.na(Charges_Joined$Desc),]
```

```{r}
Charges_Joined <- Charges_Joined %>%
  mutate(Statute = ifelse(codes == 893 & is.na(Desc),gsub("\\D{1}$", "", Statute, perl = T),
                          Statute))%>%
  select(Statute, charge, codes, Statute_orginal)
```

```{r}
Charges_Joined <- left_join(Charges_Joined, official_statues)
sum(is.na(Charges_Joined$Desc)) #66 not bad
check <- Charges_Joined[is.na(Charges_Joined$Desc),]

Charges_Joined <- select(Charges_Joined, Statute, charge, codes, Statute_orginal)
```

```{r}
#one more round of this 
Charges_Joined[Charges_Joined$Statute == 11,1] <- "1010"
Charges_Joined[Charges_Joined$Statute == '2181A4',1] <- '2181A'
Charges_Joined[Charges_Joined$Statute == 787,1] <- 64
Charges_Joined[Charges_Joined$Statute == 316611,1] <- 3160611
Charges_Joined[Charges_Joined$Statute == '3161933ABC', 1] <- '3161933ABC1'
Charges_Joined[Charges_Joined$Statute == '3161923ABC', 1] <- '3161923ABC1'
Charges_Joined[Charges_Joined$Statute == 316631, 1] <- 3160631
Charges_Joined[Charges_Joined$Statute == 316723, 1] <- 3160723
Charges_Joined[Charges_Joined$Statute == 31681, 1] <- 316801
Charges_Joined[Charges_Joined$Statute == '3161933ABC', 1] <- '3161933ABC2'
Charges_Joined[Charges_Joined$Statute == 31661, 1] <- 316610
Charges_Joined[Charges_Joined$Statute == 31629541, 1] <- 3162954
Charges_Joined[Charges_Joined$Statute == 316882, 1] <- 3160882
Charges_Joined[Charges_Joined$Statute == 316723, 1] <- 3160723
Charges_Joined[Charges_Joined$Statute == '32007003C', 1] <- '320073C'
Charges_Joined[Charges_Joined$Statute ==  32013105, 1] <- 3201315
Charges_Joined[Charges_Joined$Statute ==  '32007003B', 1] <- '320073B'
Charges_Joined[Charges_Joined$Statute ==  '32026001A', 1] <- '320261A'
Charges_Joined[Charges_Joined$Statute ==  '32007003A', 1] <- '320073A'
Charges_Joined[Charges_Joined$Statute == 324211,1] <- 3242011
Charges_Joined[Charges_Joined$Statute == '434136C',1] <- '4034136C'
Charges_Joined[Charges_Joined$Statute == 434134,1] <- 4034134
Charges_Joined[Charges_Joined$Statute == 49931,1] <- 499031
Charges_Joined[Charges_Joined$Statute == 4995012,1] <- 49900512
Charges_Joined[Charges_Joined$Statute == '78404501A2',1] <- '7840451A2'
Charges_Joined[Charges_Joined$Statute == '78404501B',1] <- '7840451B'
Charges_Joined[Charges_Joined$Statute == '78402101A',1] <- '7840211A'
Charges_Joined[Charges_Joined$Statute == '7840210L',1] <- 784021
Charges_Joined[Charges_Joined$Statute == 78404802,1] <- 7840482
Charges_Joined[Charges_Joined$Statute == '78402101A1',1] <- '7840211A'
Charges_Joined[Charges_Joined$Statute == '78404501A1A',1] <- '7840451A1'
Charges_Joined[Charges_Joined$Statute == '78404501A',1] <- '7840451A'
Charges_Joined[Charges_Joined$Statute == '78402101A',1] <- '7840211A'
Charges_Joined[Charges_Joined$Statute == '7840450L',1] <- 784045
Charges_Joined[Charges_Joined$Statute == '78404102A',1] <- '7840412A'
Charges_Joined[Charges_Joined$Statute ==  795301,1] <- 7900531
Charges_Joined[Charges_Joined$Statute ==  79023301,1] <- 7902331
Charges_Joined[Charges_Joined$Statute ==  7901,1] <- 79010
Charges_Joined[Charges_Joined$Statute ==  '794114B',1] <- '7940114B'
```

```{r}
Charges_Joined <- left_join(Charges_Joined, official_statues)
sum(is.na(Charges_Joined$Desc)) #28 not bad
check <- Charges_Joined[is.na(Charges_Joined$Desc),]
```

```{r}
Charges_Joined[Charges_Joined$Statute ==  948611,1] <- 948061
Charges_Joined[Charges_Joined$Statute ==  948612,1] <- 948061
Charges_Joined[Charges_Joined$Statute ==  9444,1] <- 94440
Charges_Joined[Charges_Joined$Statute ==  943591,1] <- 9430591
Charges_Joined[Charges_Joined$Statute ==  89314704,1] <- 8931474
Charges_Joined[Charges_Joined$Statute ==  89314701,1] <- 8931471
Charges_Joined[Charges_Joined$Statute ==  '89313501',1] <- '8931351F'
Charges_Joined[Charges_Joined$Statute ==  '89313501A1',1] <- '8931351A1'
Charges_Joined[Charges_Joined$Statute ==  '89313501B1',1] <- '8931351B1'
Charges_Joined[Charges_Joined$Statute ==  '89313501B1',1] <- '8931351B1'
Charges_Joined[Charges_Joined$Statute ==  '89313501K2',1] <- '8931351K2'
Charges_Joined[Charges_Joined$Statute ==  '89313501K2',1] <- '8931351K2'
Charges_Joined[Charges_Joined$Statute ==  '89313501C1',1] <- '8931351C1C'
Charges_Joined[Charges_Joined$Statute ==   89314701,1] <- 8931471
Charges_Joined[Charges_Joined$Statute ==   '893136A2',1] <- '893136A'
Charges_Joined[Charges_Joined$Statute ==   '838211A',1] <- '8380211A'
Charges_Joined[Charges_Joined$Statute ==   '83102B',1] <- 83102
Charges_Joined[Charges_Joined$Statute ==   831322,1] <- 8310322
Charges_Joined[Charges_Joined$Statute ==   '812132BM',1] <- '812132B'
Charges_Joined[Charges_Joined$Statute ==   '782711A',1] <- '7820711A'
Charges_Joined[Charges_Joined$Statute ==   '782711A',1] <- '7820711A'
Charges_Joined[Charges_Joined$Statute ==   '777031C2A',1] <- '777032A'
Charges_Joined[Charges_Joined$Statute ==   '777031C2A',1] <- '777032A'

Charges_Joined <- Charges_Joined %>% select(Statute, charge, codes, Statute_orginal)
```

```{r}
Charges_Joined <- left_join(Charges_Joined, official_statues)
sum(is.na(Charges_Joined$Desc)) #7 not bad
check <- Charges_Joined[is.na(Charges_Joined$Desc),]
```

```{r}
Charges_Joined[Charges_Joined$Statute == 320216,1] <- 320261
Charges_Joined[Charges_Joined$Statute == 32203101,1] <- 3220311
Charges_Joined[Charges_Joined$Statute == '32231B',1] <- '322031B'
Charges_Joined[Charges_Joined$Statute == '322516A',1] <- '3220516A'

Charges_Joined <- Charges_Joined %>% select(Statute, charge, codes, Statute_orginal)
```

```{r}
Charges_Joined <- left_join(Charges_Joined, official_statues)
sum(is.na(Charges_Joined$Desc))
check <- Charges_Joined[is.na(Charges_Joined$Desc),]

## cannot find the matching statutes in the official statutes for the following: 
    # BUSINESS W/O LIC
    # LSC/INJURY/7-1-14
    # LSC/INJURY/AT/7/1/14
```

```{r}
rm(check)
Charges_Joined <- Charges_Joined %>%
  distinct()
```

```{r}
Charges <- Charges_Joined %>%
  group_by(codes, Type) %>%
  summarise(Category = paste(charge, collapse = ";"))
```

# Count Number of Charges 

```{r}
# create variable charge_count will tell how many charges each defendant has 

books_fil <- books_fil %>%
  mutate(id = row_number()) %>%
  group_by(id) %>%
  mutate(charge_count = sum(str_detect(Charge2,""), str_detect(Charge1,""), str_detect(Charge3, "")))
```


# Create Categories more in depth

## 322 and 324 codes DWLS 

```{r}
codes_322_fel <- unlist(str_split(Charges$Category[21], pattern = ";"))
codes_322_fel <- codes_322_fel[!duplicated(codes_322_fel)]

codes_322_notfel <- unlist(str_split(Charges$Category[22:26], pattern = ";"))
codes_322_notfel <- codes_322_notfel[!duplicated(codes_322_notfel)]

codes_322_all <- c(codes_322_fel, codes_322_notfel)
```

```{r}
# Those with 322 or 324 codes either 1 and the rest are blank, 2 and the rest are blank, or all 3 have these codes. Some may be felonies, but they are all related to DWLS so I'll count them. 

books_fil <- books_fil %>% mutate(DWLS_ONLY = case_when(
  Charge1 %in% codes_322_all & Charge2 %in% codes_322_all & Charge3 %in% codes_322_all ~ 1,
  Charge1 %in% codes_322_all & Charge2 %in% codes_322_all & Charge3 == "" ~ 1,
  Charge1 %in% codes_322_all & Charge2 == "" & Charge3 == "" ~1,
  Charge1 == "" & Charge2 %in% codes_322_all & Charge3 %in% codes_322_all ~1,
  Charge1 == "" & Charge2 %in% codes_322_all & Charge3 == "" ~1,
  Charge1 == "" & Charge2 == "" & Charge3 %in% codes_322_all ~1,
  !Charge1 %in% codes_322_all | !Charge2 %in% codes_322_all | !Charge3 %in% codes_322_all ~0))

# DWLS - Felony charge 
books_fil <- books_fil  %>% mutate(DWLS_FELONY = ifelse(Charge1 %in% codes_322_fel | Charge2 %in% codes_322_fel | Charge3 %in% codes_322_fel, 1,0))
```

## 322 and 324 codes with probation, felony, alias capias, bench warrant, etc..relevant charges 

```{r}
other_charges <- c("BENCH WARRANT", "OUT-OF-CNTY/WARRANT", "ALIAS CAPIAS", "PROBATION WARRANT", "ARREST WARRANT", "WRIT/BODILY ATTACH", "PROBATION VIOLATION", "OUT ON FELONY BOND", "PRETRIAL REL/VIOLATE", "BOND SURRENDER","HOLD FOR MAGISTRATE", "OUT ON PROBATION", "CAPIAS", "WRIT OF ATTACHMENT")

books_fil <- books_fil %>% mutate(DWLS_HIGH = case_when(
  Charge1 %in% codes_322_all & Charge2 %in% codes_322_all & Charge3 %in% codes_322_all ~ 1,
  Charge1 %in% codes_322_all & Charge2 %in% codes_322_all & Charge3 == "" ~ 1,
  Charge1 %in% codes_322_all & Charge2 == "" & Charge3 == "" ~1,
  Charge1 == "" & Charge2 %in% codes_322_all & Charge3 %in% codes_322_all ~1,
  Charge1 == "" & Charge2 %in% codes_322_all & Charge3 == "" ~1,
  Charge1 == "" & Charge2 == "" & Charge3 %in% codes_322_all ~1,
  Charge1 %in% codes_322_all & Charge2 %in% other_charges & Charge3 %in% other_charges~1,
  Charge1 %in% codes_322_all & Charge2 %in% other_charges & Charge3 == "" ~1,
  Charge1 %in% codes_322_all & Charge2 %in% other_charges & Charge3 %in% codes_322_all ~1,
  Charge1 %in% codes_322_all & Charge2 %in% codes_322_all & Charge3 %in% other_charges ~1,
  Charge1 %in% other_charges & Charge2 %in% codes_322_all & Charge3 %in% codes_322_all ~1,
  Charge1 %in% other_charges & Charge2 %in% codes_322_all & Charge3 == "" ~1,
  Charge1 %in% other_charges & Charge2 %in% other_charges & Charge3 %in% codes_322_all ~1,
  Charge1 %in% other_charges & Charge2 %in% codes_322_all & Charge3 %in% other_charges ~1)) %>%
  mutate(DWLS_HIGH = ifelse(is.na(DWLS_HIGH),0, DWLS_HIGH))
```

## 316-320 Traffic/ Driving Statutes

```{r}
#316 felony charges 
codes_316_fel <- unlist(str_split(Charges$Category[9], pattern = ";"))
codes_316_fel <- codes_316_fel[!duplicated(codes_316_fel)]
codes_316_fel <- c(codes_316_fel, "LSC/INJURY/7-1-14", "LSC/INJURY/AT/7/1/14")

#319 felony charges 
codes_319_fel <- unlist(str_split(Charges$Category[16], pattern = ";"))
codes_319_fel <- codes_319_fel[!duplicated(codes_319_fel)]

#320 felony charges - there is only one 
codes_320_fel <- unlist(str_split(Charges$Category[17], pattern = ";"))
codes_320_fel <- codes_320_fel[!duplicated(codes_320_fel)]

# 316 FELONY CHARGES: 
print(codes_316_fel)

# 319 FELONY CHARGES:
print(codes_319_fel)

# 320 FELONY CHARGES: 
print(codes_320_fel)

# All felony charges
driving_traffic_fel <- c(codes_316_fel, codes_319_fel, codes_320_fel)

# All driving/traffic felony charges 
print(driving_traffic_fel)
```

```{r}
# 316 infractions
codes_316_inf <- unlist(str_split(Charges$Category[10], pattern = ";"))
codes_316_inf <- codes_316_inf[!duplicated(codes_316_inf)]

# 320 infractions
codes_320_inf <- unlist(str_split(Charges$Category[18], pattern = ";"))
codes_320_inf <- codes_320_inf[!duplicated(codes_320_inf)]

# 316 INFRACTION CHARGES: 
print(codes_316_inf)

# 320 Infraction charges
print(codes_320_inf)

# All infraction charges
driving_traffic_inf <- c(codes_316_inf, codes_320_inf)

# All driving/traffic infractions 
print(driving_traffic_inf)
```

```{r}
# 316 misdeamnors
codes_316_misd <- unlist(str_split(Charges$Category[11], pattern = ";"))
codes_316_misd <- codes_316_misd[!duplicated(codes_316_misd)]

# 318 misdeamnors
codes_318_misd <- unlist(str_split(Charges$Category[14], pattern = ";"))
codes_318_misd <- codes_318_misd[!duplicated(codes_318_misd)]

# 320 misdeamnors 
codes_320_misd <- unlist(str_split(Charges$Category[19], pattern = ";"))
codes_320_misd <- codes_320_misd[!duplicated(codes_320_misd)]

# 316 misd 
print(codes_316_misd)

# 318 misd
print(codes_318_misd)

# 320 misd 
print(codes_320_misd)

# All driving/ traffic misd. 
driving_traffic_misd <- c(codes_316_misd, codes_318_misd, codes_320_misd)
driving_traffic_misd <- driving_traffic_misd[!duplicated(driving_traffic_misd)]
```

```{r}
# 316 traffic 
codes_316_traff <- unlist(str_split(Charges$Category[12], pattern = ";"))
codes_316_traff <- codes_316_traff[!duplicated(codes_316_traff)]

# 318 traffic 
codes_318_traff <- unlist(str_split(Charges$Category[15], pattern = ";"))
codes_318_traff <- codes_318_traff[!duplicated(codes_318_traff)]

# 320 traffic 
codes_320_traff <- unlist(str_split(Charges$Category[20], pattern = ";"))
codes_320_traff <- codes_320_traff[!duplicated(codes_320_traff)]

# all driving/traffic traffic offenses 
driving_traffic_traff <- c(codes_316_traff, codes_318_traff, codes_320_traff)
driving_traffic_traff <- driving_traffic_traff[!duplicated(driving_traffic_traff)]
```

```{r}
# so we have driving_traffic_fel, which are all felony charges

# then we have driving_traffic_inf, driving_traffic_misd, driving_traffic_traff, which are minor offesnes - will collapse these three categories 
driving_traffic_ns <- c(driving_traffic_inf, driving_traffic_misd, driving_traffic_traff)
driving_traffic_ns <- driving_traffic_ns[!duplicated(driving_traffic_ns)]

# all felony + ns
driving_traffic_all <- c(driving_traffic_fel, driving_traffic_ns)
rm(driving_traffic_inf, driving_traffic_misd, driving_traffic_traff)

print(driving_traffic_fel) #felonies
print(driving_traffic_ns) # infractions, misdeamnors, traffic
```


```{r}
# create three groups -- Variable name (Serious_Driving_Traffic)
    # (1) Serious_Driving_Traffic == "S", which corresponds to 316-320 felony charges (driving_traffic_fel)
    # (2) Serious_Driving_Traffic == "NS", which corresponds to those that can be charged as misdemeanor, traffic, or infraction
                          # (driving_traffic_ns)
    # (3) Serious_Driving_Traffic == "None", those who do not have any traffic charges

books_fil <- books_fil %>%
  mutate(Serious_Driving_Traffic = case_when(
    Charge1 %in% driving_traffic_fel & Charge2 %in% driving_traffic_fel & Charge3 %in% driving_traffic_fel ~  'S',
    Charge1 %in% driving_traffic_fel & !Charge2 %in% driving_traffic_fel & !Charge3 %in% driving_traffic_fel ~ 'S',
    Charge1 %in% driving_traffic_fel & !Charge2 %in% driving_traffic_fel & Charge3 %in% driving_traffic_fel ~ 'S',
    Charge1 %in% driving_traffic_fel & Charge2 %in% driving_traffic_fel & !Charge3 %in% driving_traffic_fel ~ 'S',
    !Charge1 %in% driving_traffic_fel & Charge2 %in% driving_traffic_fel & Charge3 %in% driving_traffic_fel ~  'S',
    !Charge1 %in% driving_traffic_fel & Charge2 %in% driving_traffic_fel & !Charge3 %in% driving_traffic_fel ~ 'S',
    !Charge1 %in% driving_traffic_fel & !Charge2 %in% driving_traffic_fel & Charge3 %in% driving_traffic_fel ~ 'S', 
    Charge1 %in% driving_traffic_ns & Charge2 %in% driving_traffic_ns & Charge3 %in% driving_traffic_ns ~ 'NS',
    Charge1 %in% driving_traffic_ns & !Charge2 %in% driving_traffic_fel & !Charge3 %in% driving_traffic_fel ~ 'NS',
    !Charge1 %in% driving_traffic_fel & Charge2 %in% driving_traffic_ns & !Charge3 %in% driving_traffic_fel ~ 'NS',
    !Charge1 %in% driving_traffic_fel & !Charge2 %in%driving_traffic_fel & Charge3 %in% driving_traffic_ns ~ 'NS' , 
    !Charge1 %in% driving_traffic_all & !Charge2 %in% driving_traffic_all & !Charge3 %in% driving_traffic_all ~ 'None'
  ))
```

## County and Municipal Ordinances / Noncriminal (Minor offenses)

```{r}
# County Ordinance - C
# Municipal Ordinance - L
# Noncriminal - N

# county ordinances
county_ord <-unlist(str_split(Charges$Category[Charges$Type == "C"], pattern = ";"))
county_ord <- county_ord[!duplicated(county_ord)] 
county_ord <- county_ord[!is.na(county_ord)]

# municipal ordinances 
mun_ord <- unlist(str_split(Charges$Category[Charges$Type == "L"], pattern = ";"))
mun_ord <- mun_ord[!duplicated(mun_ord)]
mun_ord <- mun_ord[!is.na(mun_ord)]

# non criminal
noncrim <- unlist(str_split(Charges$Category[Charges$Type == "N"], pattern = ";"))  
noncrim <- noncrim[!duplicated(noncrim)]
noncrim <- noncrim[!is.na(noncrim)]

minor_offenses <- c(county_ord, mun_ord, noncrim)
minor_offenses <- minor_offenses[!duplicated(minor_offenses)]

print(minor_offenses)
```


```{r}
# Create minor offense dummy variable: Minor_Offense--will indicate whether one of the charges is a minor_offense
# Minor_offense == 1, one of the charges is a minor offense (municipal, county, or noncriminal)
# Minor_offense == 0, one of the charge is not a minor offense 

books_fil <- books_fil %>%
  group_by(id) %>%
  mutate(minor_offense = ifelse(Charge1 %in% minor_offenses | Charge2 %in% minor_offenses | Charge3 %in%
                                  minor_offenses,1,0))
```

## Major Felony charges, not related to traffic or driving 

```{r}
# statute codes - 403, 499, 775, 777, 782, 784, 787, 790, 794, 796, 806, 810, 812, 817, 827, 828, 831, 838, 843, 859, 893, 901, 914, 918, 934, 943, 944, 951 

# major felonies
major_felonies <- Charges %>%
  filter(Type == "F" & codes %in% c(403, 499, 775, 777, 782, 784, 787, 790, 794, 796, 806, 810, 812, 817, 827, 828, 831, 838, 843, 859, 893, 901, 914, 918, 934, 943, 944, 951))

major_felonies <- unlist(str_split(major_felonies$Category, pattern = ";"))
major_felonies <- major_felonies[!duplicated(major_felonies)]

print(major_felonies)
```

```{r}
# create major felony indicator -- Major_Felony will indicate whether one of the defendant's charge is a major_felony 
    # Major_Felony == 1; has a major felony charge 
    # Major_Felony == 0; does not have a major felony charge 

books_fil <- books_fil %>%
  group_by(id) %>%
  mutate(major_felony = ifelse(Charge1 %in% major_felonies | Charge2 %in% major_felonies | Charge3 %in%
                                  major_felonies,1,0))
```

# Analysis without those with major felonies

```{r}
books_fil %>%
  group_by(year) %>%
  filter(major_felony == 0) %>%
  summarise(n=n(),FTP.Ratio = ceiling((n*.05)), one.day.cost = 230*FTP.Ratio,
  two.day.cost = (2*230)*FTP.Ratio, average.stay= (30*230)*FTP.Ratio)
```































































# OLD CODE, BUT USEFUL

```{r}
# books_fil <- books_fil %>% mutate(DWLS = case_when(
#   Charge1 %in% rel_charges & Charge2 %in% rel_charges & Charge3 %in% rel_charges ~ 1,
#   Charge1 %in% rel_charges & Charge2 %in% rel_charges & Charge3 == "" ~1,
#   Charge1 %in% rel_charges & Charge2 == "" & Charge3 %in% rel_charges ~1,
#   Charge1 %in% rel_charges & Charge2 == "" & Charge3 == "" ~1,
#   Charge1 %in% rel_charges & Charge2 %in% rel_charges & Charge3 %in% other_charges ~1,
#   Charge1 %in% rel_charges & Charge2 %in% other_charges & Charge3 %in% other_charges~1,
#   Charge1 %in% rel_charges & Charge2 %in% other_charges & Charge3 == "" ~1,
#   Charge1 %in% rel_charges & Charge2 %in% other_charges & Charge3 %in% rel_charges ~1,
#   Charge1 == "" & Charge2 %in% rel_charges & Charge3 == "" ~1,
#   Charge1 == "" & Charge2 == "" & Charge3 %in% rel_charges ~ 1,
#   Charge1 %in% other_charges & Charge2 %in% rel_charges & Charge3 %in% rel_charges ~1,
#   Charge1 %in% other_charges & Charge2 %in% rel_charges & Charge3 == "" ~1,
#   Charge1 %in% other_charges & Charge2 %in% other_charges & Charge3 %in% rel_charges ~1,
#   Charge1 %in% other_charges & Charge2 %in% rel_charges & Charge3 %in% other_charges ~1,
#   !Charge1 %in% rel_charges|!Charge1 %in% other_charges ~0 ))
```

```{r}
 # DWLS_Likely <- books_fil %>% filter(DWLS ==1)
```